sem

### 2022-12-12 scripts/workshop/sem.Rmd
add lagged data

```{r}
prev_mod <- dagitty("dag{
  temp -> cc
  temp -> size
  cc -> prev
  size -> prev
  temp -> prev
             }")

@@

@@

cc ~ mean water temp; size ~ mean water temp; prevalence ~ cc + size + mean water temp
```{r}
cc_temp <- glm(log10(cc) ~ (MeanWaterTempPredC), data = comm_data)
size_temp <- glm(log10(size) ~ (MeanWaterTempPredC) + log10(cc), data = comm_data)
prev_cc <- glm(Prevalence ~ log10(cc) + log10(size), data = comm_data)

sem <- psem(cc_temp, size_temp, prev_cc)
```

@@

@@

>>>>

Make lagged data
```{r}
prev_cc <- comm_data

order <- prev_cc[order(prev_cc$WetAltID, prev_cc$Month.1),]

#this removes months out of sequence
order <- order[-c(4,14,47,90),]

order %<>% add_column(lag_cc = NA, lag_size = NA, lag_temp = NA)

#create new column that includes previous month's value for cc
for(n in 2:nrow(order)){
  order$lag_cc[n] <- order$cc[n-1]
  order$lag_size[n] <- order$size[n-1]
  order$lag_temp[n] <- order$MeanWaterTempPredC[n-1]
}

#remove the first entry for each wetland to remove the carryover from the last wetland
clean <- order %>% group_by(WetAltID) %>% filter(duplicated(WetAltID) | n()==1)

```{r}
prev_mod <- dagitty("dag{
  temp -> cc
  temp -> size
  temp -> prev
  cc -> prev
  size -> prev
             }")

@@

Covariance-based SEM
```{r}

@@

cc ~ mean water temp; size ~ mean water temp; prevalence ~ cc + size + mean water temp
```{r}
cc_temp <- lm(log10(cc) ~ (MeanWaterTempPredC), data = comm_data)
size_temp <- lm(log10(size) ~ (MeanWaterTempPredC), data = comm_data)
prev_cc <- lm(Prevalence ~ log10(cc) + log10(size), data = comm_data)
prev_temp_cc <- glm(Prevalence ~ log10(cc) + log10(size) + MeanWaterTempPredC, data = comm_data)

sem <- psem(cc_temp, size_temp, prev_cc)
sem_1 <- psem(cc_temp, size_temp, prev_temp_cc)
```

Using lagged data
```{r}
lagged_cc_temp <- lm(lag_cc ~ lag_temp, data = clean)
lagged_size_temp <- lm(log10(lag_size) ~ lag_temp + lag_cc, data = clean)
lagged_prev_cc <- lm(Prevalence ~ lag_cc + log10(lag_size), data = clean)
lagged_prev_temp_cc <- glm(Prevalence ~ lag_cc + log10(lag_size) + lag_temp, data = clean)

sem_lagged <- psem(lagged_cc_temp, lagged_size_temp, lagged_prev_cc)
sem_1_lagged <- psem(lagged_cc_temp, lagged_size_temp, lagged_prev_temp_cc)
```

@@

```{r}
basisSet(sem_1)
dSep(sem_1, conditioning = T)
fisherC(sem_1)
coefs(sem_1, intercepts = T)
rsquared(sem_1)
```

@@

```{r}
summary(sem_1)
plot(sem_1)
plot(sem_1, return = T)
```

```{r}
summary(sem_lagged)
plot(sem_lagged)
plot(sem_lagged, return = T)
```
